DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car

### 2017-10-03 paper/evaluation.tex
Add line breaks
evaluation revision subsection A.

% 50-200Hz for quadcopters:
% https://robotics.stackexchange.com/questions/231/what-frequency-does-my-quadcopter-output-sense-calculate-output-update-loop-need

@@

Figure~\ref{fig:control-loop-timing} shows the measured control loop
processing times of the DeepPicar in which 1001 image frames, from the
camera, are fed to the DNN, one frame per each control loop. The
measured time is total execution time of a control loop.

%% In evaluating the real-time efficacy of the DeepPicar, the methodology is consistent across all 
%% experiments. The performance of the platform is measured over a set of 1001 video frames that are 
%% each individually fed to the model. The processing time for the first frame is omitted as, due to 
%% cache warmup, it is uncharacteristically high and doesn't accurately represent the Picar's capabilities. 
%% A WCET of 50 ms, or 20 Hz, is used as a baseline to assess the platform's ability to complete all 
%% necessary real-time operations in a timely manner.

We seek to determine if the DeepPicar is capable of consistently executing all necessary functions 
before their given deadlines. In the case of our platform, it has to process every given frame and get 
the predicted angles from the model within a WCET of 50 ms. In order to determine if this was achievable, 
we tested the platform by running a single model that utilized all 4 cpu cores and measured the time it 
took for each frame to be processed. The performance of the DeepPicar can be seen in Figure 6. We found 
that DeepPicar was able to completely process the vast majority of the provided frames within 50 ms, and 
was unable to do so for very few frames.

In our platform, three main real-time tasks are performed during autonomous operation. In order, these 
operations are: (1) capturing and reading the input frame from the designated camera or video stream, 
(2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) feeding the frame 
to, and getting the angle prediction from, the model. We aim to determine which operation(s), if any, 
require the most time to be performed.

\begin{table}[h]
  \centering
  \begin{tabular} {| l | l | l | l | l |}
    \hline
    \textbf{operation} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
    Frame Capture & 2.28 & 4.94 & 4.54  & 0.52\\ \hline
    Preprocessing & 3.09 & 4.60 & 3.31 & 0.10 \\ \hline
    Angle Prediction & 37.30 & 51.03 & 45.48 & 2.75 \\  \hline
    Total Time & 42.67 & 56.37 & 50.70 & 2.80 \\
    \hline
  \end{tabular}
  \caption{Real-time performance of DeepPicar for each autonomous vehicle operation performed.}
\end{table}

In order to determine which operation(s) take the longest to execute, we measured the time it 
took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
and only one model was run. As is shown in Table II, the angle prediction operation consumes the 
majority of the processing for each frame. Furthermore, the time it takes for the operation to 
complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
other hand, both the frame capture and preprocessing operations take substantially less time 
and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.

>>>>

% 50-200Hz for quadcopters:
% https://robotics.stackexchange.com/questions/231/what-frequency-does-my-quadcopter-output-sense-calculate-output-update-loop-need 
% https://quadmeup.com/pid-looptime-why-it-is-not-only-about-frequency/

@@

\begin{table}[t]
  \centering
  \begin{tabular} {| c | r | r | r | r |}
    \hline
    \textbf{Operation} & \textbf{Mean} & \textbf{Max} &   \textbf{99pct.} & \textbf{Stdev.} \\ \hline
    Image capture        & 2.28  &  4.94 &  4.54  & 0.52 \\ \hline
    Image pre-processing & 3.09  &  4.60 &  3.31  & 0.10 \\ \hline
    DNN inferencing      & 37.30 & 51.03 & 45.48  & 2.75 \\ \hline
    Total Time           & 42.67 & 56.37 & 50.70  & 2.80 \\ \hline
  \end{tabular}
  \caption{Control loop timing breakdown.}
  \label{tbl:control-loop-breakdown}
\end{table}

Figure~\ref{fig:control-loop-timing} shows the measured control loop 
processing times of the DeepPicar over 1000 image frames (one per each
control loop). We omit the first frame processing time for cache
warmup. Table~\ref{tbl:control-loop-breakdown} shows the time
breakdown of each control loop. Note that all four CPU cores of the
Raspberry Pi 3 were used by the Tensorflow library when performing the
DNN inference operations.

First, as expected, we find that the inference operation
dominates the control loop execution time, accounting about 85\% of
the execution time.
Second, more importantly, we also find that the measured average
execution time of a single control loop is 42.67 ms, or 23.4 Hz and
the 99 percentail time is 50.70ms.
This means that the DeepPicar can operates
at about 20 Hz control frequency in real-time using only the on-board
Raspberry Pi 3 computing platform, without needing any remote computing
resources. We consider these results respectable given the complexity
of the deep neural network  and the fact that the inference operation
performed by Tensorflow only utilizes the CPU cores of the
Raspberry Pi 3 (its GPU is not supported by Tensorflow.)

In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
network architecture, is reportedly run at 30 Hz~\cite{Bojarski2016},
just a bit faster than the DeepPicar. Although we believe it was not
limited by their computing platform (we will experimentally compare
performance differences among multiple embedded computing platforms,
including NVIDIA's Jetson TK2, later in
Section~\ref{sec:eval-platforms}), the fact that the low-cost
Raspberry Pi 3 can achieve similar real-time control performance is
surprising.

%% In our platform, three main real-time tasks are performed during autonomous operation. In order, these 
%% operations are: (1) capturing and reading the input frame from the designated camera or video stream, 
%% (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) feeding the frame 
%% to, and getting the angle prediction from, the model. We aim to determine which operation(s), if any, 
%% require the most time to be performed.

%% In order to determine which operation(s) take the longest to execute, we measured the time it 
%% took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
%% and only one model was run. As is shown in Table II, the angle prediction operation consumes the 
%% majority of the processing for each frame. Furthermore, the time it takes for the operation to 
%% complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
%% other hand, both the frame capture and preprocessing operations take substantially less time 
%% and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.

%% In evaluating the real-time efficacy of the DeepPicar, the methodology is consistent across all 
%% experiments. The performance of the platform is measured over a set of 1001 video frames that are 
%% each individually fed to the model. The processing time for the first frame is omitted as, due to 
%% cache warmup, it is uncharacteristically high and doesn't
%% accurately represent the Picar's capabilities.  
%% A WCET of 50 ms, or 20 Hz, is used as a baseline to assess the platform's ability to complete all 
%% necessary real-time operations in a timely manner.

%% We seek to determine if the DeepPicar is capable of consistently executing all necessary functions 
%% before their given deadlines. In the case of our platform, it has to process every given frame and get 
%% the predicted angles from the model within a WCET of 50 ms. In order to determine if this was achievable, 
%% we tested the platform by running a single model that utilized all 4 cpu cores and measured the time it 
%% took for each frame to be processed. The performance of the DeepPicar can be seen in Figure 6. We found 
%% that DeepPicar was able to completely process the vast majority of the provided frames within 50 ms, and 
%% was unable to do so for very few frames.

%% In our platform, three main real-time tasks are performed during autonomous operation. In order, these 
%% operations are: (1) capturing and reading the input frame from the designated camera or video stream, 
%% (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) feeding the frame 
%% to, and getting the angle prediction from, the model. We aim to determine which operation(s), if any, 
%% require the most time to be performed.

%% In order to determine which operation(s) take the longest to execute, we measured the time it 
%% took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
%% and only one model was run. As is shown in Table II, the angle prediction operation consumes the 
%% majority of the processing for each frame. Furthermore, the time it takes for the operation to 
%% complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
%% other hand, both the frame capture and preprocessing operations take substantially less time 
%% and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.
